\LARGE \bf Towards a Data-driven Approach to Human Preferences in Motion Planning

### 2015-02-28 paper2015asml.tex
adding movie

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{wrapfig}
\usepackage{graphicx,color,soul}
\usepackage{adjustbox}

@@

\title{\LARGE \bf
Towards a Data-driven Approach to Motion Planning
}

\author{Arjun Menon$^{1}$, Pooja Kacker$^{2}$, Sachin Chitta$^{1}$% <-this % stops a space
\thanks{*This work was partly supported by NSF NRI Grant IIS 1317849}% <-this % stops a space
\thanks{$^{1}$SRI International, Menlo Park, CA
        {\tt\small arjun.menon@sri.com, sachin.chitta@sri.com}}%
\thanks{$^{2}$Independent Consultant, Menlo Park, CA
        {\tt\small }}%
}

@@

We extended the MoveIt!~\cite{MoveIt} benchmarking architecture~\cite{cohen2012generic} with video generation capabilities to automate the generation of a large number of trajectories. The MoveIt! architecture allows for the definition of environments, referred to as planning scenes, start and goal states for robots (motion planning queries). It also allows for the use of any supported planning algorithms (planners). A combination of different scenes, motion queries, and planners was used to generate robot trajectories for evalutation. We chose three distinct scenes (Fig.~\ref{fig:scenes}): a kitchen (Countertop), an industrial workbench (Industrial) and a drawer on a table (Drawer). Multiple queries were generated in each scene for trajectories of the right arm of the PR2 robot (Fig.~\ref{fig:PR2}. The robot was positioned in the scenes manually and motion planning queries were constructed by hand. The positioning in the scenes was done with the expectation that the planning queries would not be impossible or trivial for the planning algorithms to solve.

@@

The planning algorithms that we used are RRT-Connect~\cite{kuffner2000rrt} and RRT*.  RRT*~\cite{frazzoli-RRTstar} was used with two different cost functions: PathLength and MaxMinClearance.The objective for PathLength minimizes the length of the trajectory (as the sum of the euclidean distances between ends of trajectory segments). The objective for MaxMinClearance computes uses the minimum clearance of a configuration belonging to a trajectory as the value it is trying to maximize. Both planners are implemented in the OMPL library~\cite{OMPL:2012}. All the paths generated by RRT-Connect are shortcutted for a finite amount of time to improve the quality of the paths. The planned trajectories are post-processed by an iterative-time parameterization method that time-parameterizes the trajectories while taking into account velocity and accelerations. The infrastructure we designed around MoveIt! allows us to automatically generate movies of the generated trajectories using the ROS Visualizer (Rviz). We generated movies from different viewpoints to maximize the visibility of the motion to human observers reviewing it. A total of 103 different trajectories were generated in this manner. Note that although the planners we use are randomized, the post-processing step (short-cutting) helps significantly in smoothing the plans.

@@

\section{Survey Results and Analysis}
\label{sec:raw_results}
The survey resulted in at least 20 ratings for each trajectory (so there were a total of 2060 ratings from participants in the survey). In this section, we will first analyze the raw results from the survey (which are interesting in themselves). Subsequently, we will present a set of features that were used to represent individual trajectories. These features will be used in the next section to learn the human-observer ratings for robot trajectories. 
\subsection{Raw Survey Results}
\begin{figure}[h]
  \centering{\includegraphics[width=0.45\linewidth]{pictures/efficiency}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/elegant}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/smooth}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/overall}}
  \caption{Histogram of averaged ratings for the 4 different attributes for 103 trajectories. Clockwise from top left: Efficient vs. Inefficient, Elegant vs. Awkward, Smooth vs. Rough, Overall Rating}
\label{fig:survey_raw}
\end{figure}
The raw results of the survey are summarized in Fig.~\ref{fig:survey_raw} which shows a histogram of averaged responses for each of the four survey questions across the 103 trajectories that were rated by users. It is clear that a majority of the trajectories in our survey were well received by the human observers but there were quite a few trajectories that were rated badly as well. An analysis of the ratings for each trajectory showed that four (averaged) ratings for each individual trajectory were always similar to each other, indicating that users had similar preferences for each trajectory regardless of the individual attributes used to describe the preference for the trajectory (efficient, elegant, smooth, overall). Table~\ref{tab:correlation} shows the Pearson's correlation coefficients corresponding to pairs of attributes. Attributes across individual questions are highly correlated at a statistically significant level of 1\% (with p-values $< 0.01\%$). The Spearman rank correlation coefficients are also high and statistically significant. 
\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
& Efficient & Elegant & Smooth & Overall \\ \hline
Efficient & 1 & 0.96 & 0.91 & 0.98 \\ \hline
Elegant & 0.96 & 1 & 0.96 & 0.98 \\ \hline
Smooth & 0.91 & 0.96 & 1 & 0.96 \\ \hline
Overall & 0.98 & 0.98 & 0.96 & 1 \\ \hline
\end{tabular}
\caption{Correlation between the four attributes in the survey.}
\label{tab:correlation}
\vspace{-0.1in}
\end{table}

We computed the correlation between observer ratings of smoothness vs. the measured smoothness of the trajectories (Table~\ref{tab:corr_smoothness}). Note that higher values of the measured smoothness feature  indicate that the trajectories are jagged or non-smooth. The correlation between the smoothness rating and the measured smoothness is not very high, indicating that human observer perception of smoothness is different from the particular measure that we used for the robot trajectories. A likely explanation is that excessive motions of the wrist (while still measured as smooth) are perceived as non-smooth by the human observer. We also computed the correlation between path lengths and the efficiency rating assigned by the human observer (Table~\ref{tab:path_length}). Here, the correlation was high (the negative value of the correlation coefficients indicate that higher path lengths are perceived as less efficient).

\begin{table}
\centering
\begin{tabular}{|p{1.25cm}|c|c|c|c|}
\hline
& Class & Smoothness & Smoothness & Smoothness \\ 
& Smooth & (Joint) & (Tool) & (Wrist) \\ \hline
Class Smooth & 1 & -0.27 & -0.29 & -0.38 \\ \hline
Smoothness (Joint) & -0.27 & 1 & 0.68 & 0.62 \\ \hline
Smoothness (Tool) & -0.29 & 0.68 & 1 & 0.87 \\ \hline
Smoothness (Wrist) & -0.38 & 0.62 & 0.87 & 1 \\ \hline
\end{tabular}
\caption{Correlation between human observed rating for smoothness and measured smoothness values for the trajectories.}
\label{tab:corr_smoothness}
\vspace{-0.1in}
\end{table}

\begin{table}
\centering
\begin{tabular}{|p{1.25cm}|c|c|c|c|}
\hline
& Class & Path Length & Path Length & Path Length \\ 
& Efficient & (Joint) & (Wrist) & (Tool) \\ \hline
\hline
Class Efficient & 1 & -0.79 & -0.8 & -0.76 \\ \hline
Path Length (Joint) & -0.79 & 1 & 0.89 & 0.86 \\ \hline
Path Length (Wrist) & -0.8 & 0.89 & 1 & 0.97 \\ \hline
Path Length (Tool) & -0.76 & 0.86 & 0.97 & 1 \\ \hline
\end{tabular}
\caption{Correlation between human observed rating for efficiency and measured path lengths for the trajectories.}
\label{tab:path_length}
\vspace{-0.1in}
\end{table}

@@

The first set of features we compute for each trajectory take into account the environment around the robot, the shape of the trajectory and the dynamics of the trajectory. These make up our unary features computed only on the input trajectory. We also include features that compare each trajectory to a nominal trajectory - these feature serve to characterize how much the trajectory deviates from the ``expected'' path. We go over these comparative features in the next section. First, we detail the features we computed on each trajectory \tj{} with waypoints $q_i$ where $i={0,\mathellipsis,n}$ and $n=|\tj{}|$. Note that our features are generic, i.e. they can be computed in the same manner for any robot arm trajectory. We believe that our methods should be easily extensible to robots other than the PR2 as well and individual trajectory classifiers could be built for each different robot arm in a similar manner.

@@

{\bf Joint Limit Distance: }
The maximin distance to joint limits is the maximum over waypoints, of the miniminum over joints of their distance to their respective closest limit (upper or lower), or $\max_{q \in \tj{}} \min_{\theta \in q} \min(|\theta-\theta_U|,|\theta-\theta_L|)$. Note that although the joint limit feature may not be immediately perceived by users, we chose to include it to see if the data might prove otherwise.

@@

{\bf Nominal Trajectory Generation}: \hl{The expected or nominal end-effector trajectories were created using a breadth-first search in a 3D grid approximation of the planning problem. We plan the path of the end-effector as if it were a }\emph{free-standing cube} \hl{that is not contrained by the kinematics of the arm to which it is attached. The environment model was first discretized into a voxel-grid, and voxels are marked as occupied or empty. Any voxel that is partially occupied is considered to be occupied.

@@

The second classification framework we use is a random forest ensemble of trees. This involves randomly dividing the data into several bootstrap training samples and building a tree on each bootstrap sample. Within each bootstrapped sample, at each split a subset of features is randomly selected from the full set of features, but only one predictor is used at each split. This brings additional randomness within each tree. The class is chosen by majority vote across all trees. For every tree constructed from a bootstrapped training sample there is one Out of Bag (OOB) error which can be thought of as a test error for every tree. It is estimated internally, during the run. On test data, the random forest method returns 83\% accuracy.

@@

\begin{figure}[t]
  \centering{\includegraphics[width=0.90\linewidth]{pictures/meandecrease}}
  \caption{Random Forest: Mean decrease in accuracy for each feature.}
\label{fig:feature_importance_1}
\vspace{-0.2in}
\end{figure}
\begin{figure}[t]
 \centering{\includegraphics[width=0.90\linewidth]{pictures/gini}}
  \caption{Random Forest: Gini index (higher values indicate more important features).}
\label{fig:feature_importance_2}
\vspace{-0.3in} 
\end{figure}
\begin{figure}[t]
 \centering{\includegraphics[width=0.90\linewidth]{pictures/gini}}
  \caption{Boosting: Gini index (higher values indicate more important features).}
\label{fig:gini_boosting}
\vspace{-0.3in}
\end{figure}

The classification results clearly show the value in using a data-driven approach to determining human preferences for robot trajectories. It is difficult, 
in obstacle filled cluttered environments, to hand-design an appropriate cost function for generating human preferred trajectories. Tuning coefficients 
and weights for all the individual cost components for designing preferred trajectories requires significant effort. 
The classifiers that we have built give us a notion of the quality of trajectories 
being generated without an explicit need to model the cost function. The output from the classifier could be integrated into gradient-free motion planning schemes 
(e.g. STOMP or a simulated annealing approach) as a cost for a particular trajectory. The identification of the key features that are important in defining a preferred trajectory 
should also help in the design of cost functions, especially since some the features may not be traditionally included as part of a cost function. 
In particular, note that a feature like wrist\_tool\_dtw is not used in typical optimization schemes like RRT* which rely on a path length criterion.

>>>>

% The following packages can be found on http:\\www.ctan.org
%\usepackage{graphics} % for pdf, bitmapped graphics files
%\usepackage{epsfig} % for postscript graphics files
%\usepackage{mathptmx} % assumes new font selection scheme installed
%\usepackage{times} % assumes new font selection scheme installed
%\usepackage{amsmath} % assumes amsmath package installed
%\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{wrapfig}
\usepackage{graphicx,color,soul}
\usepackage{adjustbox}
\usepackage{flushend}

@@

\title{\LARGE \bf
Towards a Data-driven Approach to Human Preferences in Motion Planning
}

\author{Arjun Menon$^{1}$, Pooja Kacker$^{2}$, Sachin Chitta$^{3}$% <-this % stops a space
\thanks{*This work was partly supported by NSF NRI Grant IIS 1317849}% <-this % stops a space
\thanks{$^{1}$Anki Robotics, San Francisco, CA - this work was performed while Arjun was at SRI International
        {\tt\small }}%
\thanks{$^{2}$Independent Consultant, Menlo Park, CA
        {\tt\small }}%
\thanks{$^{3}$Corresponding author: SRI International, Menlo Park, CA
        {\tt\small sachin.chitta@sri.com}}%
}

@@

We extended the MoveIt! benchmarking architecture~\cite{cohen2012generic} with video generation capabilities to automate the generation of a large number of trajectories. The MoveIt!~\cite{MoveIt}  architecture allows for the definition of environments, referred to as planning scenes, start and goal states for robots (motion planning queries). It also allows for the use of any supported planning algorithms (planners). A combination of different scenes, motion queries, and planners was used to generate robot trajectories for evalutation. We chose three distinct scenes (Fig.~\ref{fig:scenes}): a kitchen (Countertop), an industrial workbench (Industrial) and a drawer on a table (Drawer). Multiple queries were generated in each scene for trajectories of the right arm of the PR2 robot (Fig.~\ref{fig:PR2}). The robot was positioned in the scenes manually and motion planning queries were constructed by hand. The positioning in the scenes was done with the expectation that the planning queries would not be impossible or trivial for the planning algorithms to solve.

@@

The planning algorithms that we used are RRT-Connect~\cite{kuffner2000rrt} and RRT*.  RRT*~\cite{frazzoli-RRTstar} was used with two different cost functions: PathLength and MaxMinClearance.The objective for PathLength minimizes the length of the trajectory (as the sum of the euclidean distances between ends of trajectory segments). The objective for MaxMinClearance uses the minimum clearance of a configuration belonging to a trajectory as the value it is trying to maximize. Both planners are implemented in the OMPL library~\cite{OMPL:2012}. All the paths generated by RRT-Connect are shortcutted for a finite amount of time to improve the quality of the paths. The planned trajectories are post-processed by an iterative-time parameterization method that time-parameterizes the trajectories while taking into account velocity and accelerations. The infrastructure we designed around MoveIt! allows us to automatically generate movies of the generated trajectories using the ROS Visualizer (Rviz). We generated movies from different viewpoints to maximize the visibility of the motion to human observers reviewing it. A total of 103 different trajectories were generated in this manner. Note that although the planners we use are randomized, the post-processing step (short-cutting) helps significantly in smoothing the plans.

@@

\section{Survey Results and Analysis}
\label{sec:raw_results}
The survey resulted in at least 20 ratings for each trajectory (so there were a total of 2060 ratings from participants in the survey). In this section, we will first analyze the raw results from the survey (which are interesting in themselves). Subsequently, we will present a set of features that were used to represent individual trajectories. These features will be used in the next section to learn the human-observer ratings for robot trajectories. 
\subsection{Raw Survey Results}
\begin{figure}[h]
  \centering{\includegraphics[width=0.45\linewidth]{pictures/efficiency}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/elegant}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/smooth}}
  \centering{\includegraphics[width=0.45\linewidth]{pictures/overall}}
  \caption{Histogram of averaged ratings for the 4 different attributes for 103 trajectories. Clockwise from top left: Efficient vs. Inefficient, Elegant vs. Awkward, Smooth vs. Rough, Overall Rating}
\label{fig:survey_raw}
\end{figure}
The raw results of the survey are summarized in Fig.~\ref{fig:survey_raw} which shows a histogram of averaged responses for each of the four survey questions across the 103 trajectories that were rated by users. It is clear that a majority of the trajectories in our survey were well received by the human observers but there were quite a few trajectories that were rated badly as well. An analysis of the ratings for each trajectory showed that the four (averaged) ratings for each individual trajectory were always similar to each other, indicating that users had similar preferences for each trajectory regardless of the individual attributes used to describe the preference for the trajectory (efficient, elegant, smooth, overall). Table~\ref{tab:correlation} shows the Pearson's correlation coefficients corresponding to pairs of attributes. Attributes across individual questions are highly correlated at a statistically significant level of 1\% (with p-values $< 0.01\%$). The Spearman rank correlation coefficients are also high and statistically significant. 
\begin{table}
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
& Efficient & Elegant & Smooth & Overall \\ \hline
Efficient & 1 & 0.96 & 0.91 & 0.98 \\ \hline
Elegant & 0.96 & 1 & 0.96 & 0.98 \\ \hline
Smooth & 0.91 & 0.96 & 1 & 0.96 \\ \hline
Overall & 0.98 & 0.98 & 0.96 & 1 \\ \hline
\end{tabular}
\caption{Correlation between the four attributes in the survey.}
\label{tab:correlation}
\vspace{-0.1in}
\end{table}

We computed the correlation between observer ratings of smoothness vs. the measured smoothness of the trajectories (Table~\ref{tab:corr_smoothness}). Note that higher values of the measured smoothness feature  indicate that the trajectories are jagged or non-smooth. The smoothness rating is not very highly correlated with any of the measured smoothness features.  A likely explanation is that excessive motions of the wrist (while still measured as smooth) are perceived as non-smooth by the human observer. We also computed the correlation between path lengths and the efficiency rating assigned by the human observer (Table~\ref{tab:path_length}). Here, the correlation was high (the negative value of the correlation coefficients indicate that higher path lengths are perceived as less efficient).

\begin{table}
\centering
\begin{tabular}{|p{1.25cm}|c|c|c|c|}
\hline
& Smooth & Smoothness & Smoothness & Smoothness \\ 
& & (Joint) & (Tool) & (Wrist) \\ \hline
Smooth & 1 & -0.27 & -0.29 & -0.38 \\ \hline
Smoothness (Joint) & -0.27 & 1 & 0.68 & 0.62 \\ \hline
Smoothness (Tool) & -0.29 & 0.68 & 1 & 0.87 \\ \hline
Smoothness (Wrist) & -0.38 & 0.62 & 0.87 & 1 \\ \hline
\end{tabular}
\caption{Correlation between human observed rating for smoothness and measured smoothness values for the trajectories.}
\label{tab:corr_smoothness}
\vspace{-0.1in}
\end{table}

\begin{table}
\centering
\begin{tabular}{|p{1.5cm}|c|c|c|c|}
\hline
& Efficient & Path Length & Path Length & Path Length \\ 
&  & (Joint) & (Wrist) & (Tool) \\ \hline
Efficient & 1 & -0.79 & -0.8 & -0.76 \\ \hline
Path Length (Joint) & -0.79 & 1 & 0.89 & 0.86 \\ \hline
Path Length (Wrist) & -0.8 & 0.89 & 1 & 0.97 \\ \hline
Path Length (Tool) & -0.76 & 0.86 & 0.97 & 1 \\ \hline
\end{tabular}
\caption{Correlation between human observed rating for efficiency and measured path lengths for the trajectories.}
\label{tab:path_length}
\vspace{-0.1in}
\end{table}

@@

The first set of features we compute for each trajectory take into account the environment around the robot, the shape of the trajectory and the dynamics of the trajectory. These make up our unary features computed only on the input trajectory. We also include features that compare each trajectory to a nominal trajectory - these features serve to characterize how much the trajectory deviates from the ``expected'' path. We go over these comparative features in the next section. First, we detail the features we computed on each trajectory \tj{} with waypoints $q_i$ where $i={0,\mathellipsis,n}$ and $n=|\tj{}|$. Note that our features are generic, i.e. they can be computed in the same manner for any robot arm trajectory. We believe that our methods should be easily extensible to robots other than the PR2 as well and individual trajectory classifiers could be built for each different robot arm in a similar manner.

@@

{\bf Joint Limit Distance: }
The maxmin distance to joint limits is the maximum over waypoints, of the minimum over joints of their distance to their respective closest limit (upper or lower), or $\max_{q \in \tj{}} \min_{\theta \in q} \min(|\theta-\theta_U|,|\theta-\theta_L|)$. Note that although the joint limit feature may not be immediately perceived by users, we chose to include it to see if the data might prove otherwise.

@@

{\bf Nominal Trajectory Generation}: \hl{The expected or nominal end-effector trajectories were created using a breadth-first search in a 3D grid approximation of the planning problem. We plan the path of the end-effector as if it were a }\emph{free-standing cube} \hl{that is not constrained by the kinematics of the arm to which it is attached. The environment model was first discretized into a voxel-grid, and voxels are marked as occupied or empty. Any voxel that is partially occupied is considered to be occupied.

@@

The second classification framework we use is a random forest ensemble of trees. This involves randomly dividing the data into several bootstrap training samples and building a tree on each bootstrap sample. Within each bootstrapped sample, at each split a subset of features is randomly selected from the full set of features, but only one predictor is used at each split. This brings additional randomness within each tree. The class is chosen by majority vote across all trees. For every tree constructed from a bootstrapped training sample there is one Out of Bag (OOB) error which can be thought of as a test error for every tree. Our model produced an OOB estimate of error rate equivalent to 19\%. On the test data, the random forest method returns 83\% accuracy.

@@

\begin{figure}[t]
  \centering{\includegraphics[width=0.90\linewidth]{pictures/meandecrease}}
  \caption{Random Forest: Mean decrease in accuracy for each feature.}
\label{fig:feature_importance_1}
\vspace{-0.2in}
\end{figure}
\begin{figure}[t]
 \centering{\includegraphics[width=0.90\linewidth]{pictures/gini}}
  \caption{Random Forest: Gini index based relative importance measure.}
\label{fig:feature_importance_2}
\vspace{-0.3in} 
\end{figure}
\begin{figure}[t]
 \centering{\includegraphics[width=0.90\linewidth]{pictures/gini_boosting}}
  \caption{Boosting: Gini index based relative importance measure.}
\label{fig:gini_boosting}
\vspace{-0.3in}
\end{figure}

The classification results clearly show the value in using a data-driven approach to determining human preferences for robot trajectories. It is difficult, 
in obstacle filled cluttered environments, to hand-design an appropriate cost function for generating human preferred trajectories. Tuning coefficients 
and weights for all the individual cost components for designing preferred trajectories requires significant effort. 
The classifiers that we have built give us a notion of the quality of trajectories 
being generated without an explicit need to model the cost function. The output from the classifier could be integrated into gradient-free motion planning schemes 
(e.g. STOMP or a simulated annealing approach) as a cost for a particular trajectory. The identification of the key features that are important in defining a preferred trajectory 
should also help in the design of cost functions, especially since some of the features may not be traditionally included as part of a cost function. 
In particular, note that a feature like wrist\_tool\_dtw is not used in typical optimization schemes like RRT* which rely on a path length criterion.
